### 2020-12-20 written/eLifeRevision/Bittner_eLife2020.tex
SC methods

\subsection{A deep generative modeling approach to emergent property inference} \label{results_dgm}
Emergent property inference (EPI) systematizes the three-step procedure of the previous section.
%First, we consider the model as a coupled set of differential (and potentially stochastic) equations \cite{gutierrez2013multiple}.  
First, we consider the model as a coupled set of stochastic differential equations \cite{gutierrez2013multiple}.  
In the running STG example, the model activity $\mathbf{x} = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$ is the membrane potential for each neuron, which evolves according to the biophysical conductance-based equation:
\begin{equation} 
\begin{split}
C_m \frac{d\mathbf{x}(t)}{dt} = -\mathbf{h}(\mathbf{x}(t); \mathbf{z}) = - [ &\mathbf{h}_{leak}(\mathbf{x}(t); \mathbf{z}) + \mathbf{h}_{Ca}(\mathbf{x}(t); \mathbf{z}) + \mathbf{h}_K(\mathbf{x}(t); \mathbf{z}) \\
 & + \mathbf{h}_{hyp}(\mathbf{x}(t); \mathbf{z}) + \mathbf{h}_{elec}(\mathbf{x}(t); \mathbf{z}) + \mathbf{h}_{syn}(\mathbf{x}(t); \mathbf{z})] 
 \end{split}
\end{equation} 
where $C_m$=1nF, and $\mathbf{h}_{\text{leak}}$, $\mathbf{h}_{Ca}$, $\mathbf{h}_K$, $\mathbf{h}_{\text{hyp}}$, $\mathbf{h}_{\text{elec}}$, and $\mathbf{h}_{\text{syn}}$ are the leak, calcium, potassium, hyperpolarization, electrical, and synaptic currents, all of which have their own complicated dependence on $\mathbf{x}$ and $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$ (see Section \ref{methods_STG}).

%Second, we define the emergent property, which as above is network syncing: oscillation of the entire population at an intermediate frequency of our choosing (Figure \ref{fig:STG}A bottom).
Second, we define the emergent property, which as above is middle ``hub neuron frequency": oscillation of the hub neuron at an intermediate frequency (Figure \ref{fig:STG}C).
%Quantifying this phenomenon is straightforward: we define network syncing to be that each neuron's spiking frequency -- denoted $\omega_{\text{f1}}(\mathbf{x})$, $\omega_{\text{f2}}(\mathbf{x})$, etc. -- is close to an intermediate frequency of 0.53Hz.  
%Mathematically, we achieve this via constraints on the mean and variance of $\omega_\alpha(\mathbf{x})$ for each neuron $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$:
Quantifying this phenomenon is straightforward: we stipulate that the hub neuron's spiking frequency -- denoted $\omega_{\text{hub}}(\mathbf{x})$is close to an intermediate frequency of 0.55Hz.  
Mathematically, we achieve this via constraints on the mean and variance of the hub neuron spiking frequency.

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Identifying neural mechanisms of flexible task switching} \label{results_SC}
In a rapid task switching experiment \cite{duan2015requirement}, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from a visual stimulus in the Anti (A) task (Fig. \ref{fig:SC}a). 
Neural recordings in the midbrain superior colliculus (SC) exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side): the Pro/Contra and Anti/Ipsi neurons \cite{duan2018collicular}.
Duan et al. proposed a model of SC that, like the V1 model analyzed in the previous section, is a four-population dynamical system.  
We analyzed this model, where the neuron-type populations are functionally-defined as the Pro- and Anti-populations in each hemisphere (left (L) and right (R)), their connectivity is parameterized geometrically  (Fig. \ref{fig:SC}B).
The input-output function of this model is chosen such that the population responses $x = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ are bounded from 0 to 1 giving rise to high (1) or low (0) responses at the end of the trial:
\begin{equation}
x_\alpha =\left(\frac{1}{2}\tanh\left(\frac{u_\alpha - \epsilon}{\zeta}\right)+ \frac{1}{2} \right)
\end{equation}
where $\epsilon = 0.05$ and $\zeta = 0.5$.  
The dynamics evolve with timescale $\tau=0.09$ via an internal variable $u$ governed by connectivity weights $W$
\begin{equation}
\tau \frac{du}{dt} = -u + Wx + h + \sigma dB
\end{equation}
with gaussian noise of variance $\sigma^2$ = 1.
The input $h$ is comprised of a cue-dependent input to the Pro or Anti populations, a stimulus orientation input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
Here, we use EPI to determine the changes in network connectivity $z = [ sW_P, sW_A, vW_{PA}, vW_{AP}, dW_{PA}, dW_{AP}, hW_P, hW_A]$ resulting in greater levels of rapid task switching accuracy.

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{figures/fig4/fig4.pdf}
\end{center}
\caption{\small EPI reveals changes in SC \cite{duan2018collicular} connectivity that control task accuracy.  
A. Rapid task switching behavioral paradigm (see text). 
B. Model of superior colliculus (SC). Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti. 
Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations, and e.g. $vW_{PA}$ is a vertical weight from an Anti to a Pro population.  
C. The Schur decomposition of the weight matrix $W = V\Lambda V^{-1}$ is a unique decomposition with orthogonal $V$ and upper triangular $\Lambda$. Schur modes: $v_{\text{all}}$, $v_{\text{task}}$, $v_{\text{side}}$, and $v_{\text{diag}}$.  
D. The marginal EPI distributions of the Schur eigenvalues at each level of task accuracy.
E. The correlation of Schur eigenvalue with task performance in each learned EPI distribution.}
\label{fig:SC}
\end{figure}

To quantify the emergent property of rapid task switching at various levels of accuracy, we considered the requirements of this model in this behavioral paradigm.
At the end of successful trials, the response of the Pro population in the hemisphere of the correct choice must have a value near 1, while the Pro population in the opposite hemisphere must have a value near 0.
Constraining a population response $x_\alpha \in \left[0,1\right]$ to be either 0 or 1 can be achieved by requiring that it has Bernoulli variance (see Section \ref{methods_SC}).
Thus, we can formulate rapid task switching at a level of accuracy $p \in \left[0, 1\right]$ in both tasks in terms of the average steady response of the Pro population $\hat{p}$ of the correct choice,  the error in Bernoulli variance of that Pro neuron $\sigma^2_{\text{err}}$, and the average difference in Pro neuron responses $d$ in both Pro and Anti trials:
\begin{equation}
\mathcal{B}(p) ~~\triangleq~~ \mathbb{E}\begin{bmatrix} \hat{p}_P \\ \hat{p}_A \\ (\hat{p}_P-p)^2 \\ (\hat{p}_A - p)^2 \\ \sigma^2_{P,err} \\ \sigma^2_{A,err} \\ d_P \\ d_A \end{bmatrix} = \begin{bmatrix} p \\ p \\ 0.15^2 \\ 0.15^2 \\ 0 \\ 0 \\ 1 \\ 1 \end{bmatrix}.
\end{equation}
Thus, $\mathcal{B}(p)$ denotes Bernoulli, winner-take-all responses between Pro neurons in a model executing rapid task switching near accuracy level $p$.

We used EPI to learn distributions of the SC weight matrix parameters $z$ conditioned on of various levels of rapid task switching accuracy $\mathcal{B}(p)$ for $p \in \{50\%, 60\%, 70\%, 80\%, 90\%\}$.
To make sense of these inferred distributions, we followed the approach of Duan et al. by decomposing the connectivity matrix $W = V\Lambda V^{-1}$ in such a way (the Schur decomposition) that the basis vectors $v_i$ are the same for all $W$ (Fig. \ref{fig:SC}C). 
These basis vectors have intuitive roles in processing for this task, and are accordingly named the \textit{all} mode - all neurons co-fluctuate, \textit{side} mode - one side dominates the other, \textit{task} mode - the Pro or Anti populations dominate the other, and \textit{diag} mode - Pro- and Anti-populations of opposite hemispheres dominate the opposite pair. 
The corresponding eigenvalues (e.g. $\lambda_{\text{task}}$, which change according to $W$) indicate the degree to which activity along that mode is increased or decreased by $W$.

We found that for greater task accuracies, the task mode eigenvalue increases, indicating the importance of $W$ to the task representation (Fig. \ref{fig:SC}D, purple; adjacent distributions from 60\% to 90\% have  $p<10^{-4}$, Mann-Whitney test with 50 estimates and 100 samples).
Stepping from random chance (50\%) networks to marginally task-performing (60\%) networks, there is a marked decrease of the side mode eigenvalues (Fig. \ref{fig:SC}D, orange; $p<10^{-4}$).  
Such side mode suppression relative to $50\%$ remains in the models achieving greater accuracy, revealing its importance towards task performance.
There were no interesting trends with task accuracy in the all or diag mode (hence not shown in Fig. \ref{fig:SC}). 
Importantly, we can conclude from our methodology that side mode suppression in $W$ allows rapid task switching, and that greater task-mode representations in $W$ increase accuracy.  
These hypotheses are confirmed by forward simulation of the SC model (Fig. \ref{fig:SC}E, see Section \ref{methods_SC}) suggesting experimentally testable predictions: increase in rapid task switching performance should be correlated with changes in effective connectivity corresponding to an increase in task mode and decrease in side mode eigenvalues.

@@

\subsubsection{Example: 2D LDS}\label{methods_2DLDS}
To gain intuition for EPI, consider a two-dimensional linear dynamical system (2D LDS) model (Fig. S1A):
\begin{equation} 
\tau \frac{d\mathbf{x}}{dt} = A\mathbf{x}
\end{equation}
with
\begin{equation}
A = \begin{bmatrix} a_1 & a_2 \\ a_3 & a_4 \end{bmatrix}.
\end{equation}
To run EPI with the dynamics matrix elements as the free parameters $\mathbf{z} = [ a_1, a_2, a_3, a_4]$ (fixing $\tau=1$), the emergent property statistics $T(\mathbf{x})$ were chosen to contain the first and second moments of the oscillatory frequency, $\frac{\text{imag}(\lambda_1)}{2 \pi}$, and the growth/decay factor, $\text{real}(\lambda_1)$, of the oscillating system. 
 $\lambda_1$ is the eigenvalue of greatest real part when the imaginary component is zero, and alternatively of positive imaginary component when the eigenvalues are complex conjugate pairs.  
To learn the distribution of real entries of $A$ that produce a band of oscillating systems around 1Hz, we formalized this emergent property as $\text{real}(\lambda_1)$ having mean zero with variance $0.25^2$, and the oscillation frequency $2 \pi \text{imag}(\lambda_1)$ having mean $\omega = 1$ Hz with variance (0.1Hz)$^2$:
\begin{equation}
 \mathbb{E}\left[T(\mathbf{x}) \right] ~~ \triangleq ~~ \mathbb{E} \begin{bmatrix} \text{real}(\lambda_1) \\ \text{imag}(\lambda_1) \\ (\text{real}(\lambda_1)-0)^2  \\ (\text{imag}(\lambda_1)-2 \pi \omega)^2 \end{bmatrix} = \begin{bmatrix} 0.0 \\ 2 \pi \omega \\ 0.25^2 \\ (2 \pi 0.1)^2 \end{bmatrix} ~~ \triangleq ~~ \bm{\mu}.
 \end{equation} 
\begin{figure}
\begin{center}
\includegraphics[scale=0.58]{figures/figS1/figS1.pdf}
\end{center}
\begin{flushleft}
{\small Fig. S1: A. Two-dimensional linear dynamical system model, where real entries of the dynamics matrix $A$ are the parameters.  
B. The EPI distribution for a two-dimensional linear dynamical system with $\tau=1$ that produces an average of 1Hz oscillations with some small amount of variance.  Dashed lines indicate the parameter axes. 
C. Entropy throughout the optimization.  
At the beginning of each augmented Lagrangian epoch (2,000 iterations), the entropy dipped due to the shifted optimization manifold where emergent property constraint satisfaction is increasingly weighted.  
D. Emergent property moments throughout optimization.  
At the beginning of each augmented Lagrangian epoch, the emergent property moments adjust closer to their constraints.}
\end{flushleft}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{figures/figS2/figS2.pdf}
\end{center}
\begin{flushleft}
{\small Fig. S2: A. Probability contours in the $a_1$-$a_4$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$. 
B. Probability contours in the $a_2$-$a_3$ plane were derived from the emergent property statistic of oscillation frequency $2\pi \text{imag}(\lambda_1)$.}
\end{flushleft}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{figures/figS3/figS3.pdf}
\end{center}
\begin{flushleft}
{\small Fig. S3: Sampled dynamical systems $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ and their simulated activity from $\mathbf{x}(0) = [\frac{\sqrt{2}}{2}, -\frac{\sqrt{2}}{2}]$ colored by log probability. A. Each dimension of the simulated trajectories throughout time.  
B  The simulated trajectories in phase space.}
\end{flushleft}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.6]{figures/figS4/figS4.pdf}
\end{center}
\begin{flushleft}
{\small Fig. S4: EPI optimization of the STG model producing network syncing. 
A. Entropy throughout optimization. 
B. The first moment emergent property statistics converge to the emergent property values at 10,000 iterations, following the fourth augmented Lagrangian epoch of 2,500 iterations.  
Since $q_{\bm{\theta}}(\mathbf{z})$ failed to produce enough samples yielding $\omega_{\text{f1}}(\mathbf{x})$ less than $0.53$Hz, the convergence criteria were not satisfied after the third epoch at 7,500 iterations.    
C. The second moment emergent property statistics converge to the emergent property values.}
\end{flushleft}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{figures/figS5/figS5.pdf}
\end{center}
{\small Fig. S5: A. EPI for rank-1 networks doing noisy discrimination. B. EPI for rank-2 networks doing context-dependent discrimination.  See \cite{mastrogiuseppe2018linking} for theoretical equations and task description.}
\end{figure}

>>>>

\subsection{A deep generative modeling approach to emergent property inference} \label{results_dgm}
Emergent property inference (EPI) systematizes the three-step procedure of the previous section.
%First, we consider the model as a coupled set of differential (and potentially stochastic) equations \cite{gutierrez2013multiple}.  
First, we consider the model as a coupled set of stochastic differential equations \cite{gutierrez2013multiple}.  
In the running STG example, the model activity $\mathbf{x} = \left[ x_{\text{f1}}, x_{\text{f2}}, x_{\text{hub}}, x_{\text{s1}}, x_{\text{s2}} \right]$ is the membrane potential for each neuron, which evolves according to the biophysical conductance-based equation:
\begin{equation} 
\begin{split}
C_m \frac{d\mathbf{x}(t)}{dt} = -\mathbf{h}(\mathbf{x}(t); \mathbf{z}) + d\mathbf{B}
 \end{split}
\end{equation} 
where $C_m$=1nF, and $\mathbf{h}$ is a sum of the leak, calcium, potassium, hyperpolarization, electrical, and synaptic currents, all of which have their own complicated dependence on $\mathbf{x}$ and $\mathbf{z} = [g_{\text{el}}, g_{\text{synA}}]$, and $d\mathbf{B}$ is white gaussian noise (see Section \ref{methods_STG}).

%Second, we define the emergent property, which as above is network syncing: oscillation of the entire population at an intermediate frequency of our choosing (Figure \ref{fig:STG}A bottom).
Second, we define the emergent property, which as above is ``intermediate hub  frequency" (Figure \ref{fig:STG}C).
%Quantifying this phenomenon is straightforward: we define network syncing to be that each neuron's spiking frequency -- denoted $\omega_{\text{f1}}(\mathbf{x})$, $\omega_{\text{f2}}(\mathbf{x})$, etc. -- is close to an intermediate frequency of 0.53Hz.  
%Mathematically, we achieve this via constraints on the mean and variance of $\omega_\alpha(\mathbf{x})$ for each neuron $\alpha \in \{ \text{f1}, \text{f2}, \text{hub}, \text{s1}, \text{s2} \}$:
Quantifying this phenomenon is straightforward: we stipulate that the hub neuron's spiking frequency -- denoted $\omega_{\text{hub}}(\mathbf{x})$is close to an intermediate frequency of 0.55Hz.  
Mathematically, we achieve this via constraints on the mean and variance of the hub neuron spiking frequency.

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Identifying neural mechanisms of flexible task switching} \label{results_SC}
%In a rapid task switching experiment \cite{duan2015requirement}, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from a visual stimulus in the Anti (A) task (Fig. \ref{fig:SC}a). 
%Neural recordings in the midbrain superior colliculus (SC) exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side): the Pro/Contra and Anti/Ipsi neurons \cite{duan2018collicular}.
%Duan et al. proposed a model of SC that, like the V1 model analyzed in the previous section, is a four-population dynamical system.  
%We analyzed this model, where the neuron-type populations are functionally-defined as the Pro- and Anti-populations in each hemisphere (left (L) and right (R)), their connectivity is parameterized geometrically  (Fig. \ref{fig:SC}B).
%The input-output function of this model is chosen such that the population responses $x = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ are bounded from 0 to 1 giving rise to high (1) or low (0) responses at the end of the trial:
%x_\alpha =\left(\frac{1}{2}\tanh\left(\frac{u_\alpha - \epsilon}{\zeta}\right)+ \frac{1}{2} \right)
%where $\epsilon = 0.05$ and $\zeta = 0.5$.  
%The dynamics evolve with timescale $\tau=0.09$ via an internal variable $u$ governed by connectivity weights $W$
%\tau \frac{du}{dt} = -u + Wx + h + \sigma dB
%with gaussian noise of variance $\sigma^2$ = 1.
%The input $h$ is comprised of a cue-dependent input to the Pro or Anti populations, a stimulus orientation input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
%Here, we use EPI to determine the changes in network connectivity $z = [ sW_P, sW_A, vW_{PA}, vW_{AP}, dW_{PA}, dW_{AP}, hW_P, hW_A]$ resulting in greater levels of rapid task switching accuracy.
%\includegraphics[scale=0.5]{figures/fig4/fig4.pdf}
%\caption{\small EPI reveals changes in SC \cite{duan2018collicular} connectivity that control task accuracy.  
%A. Rapid task switching behavioral paradigm (see text). 
%B. Model of superior colliculus (SC). Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti. 
%Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
%Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations, and e.g. $vW_{PA}$ is a vertical weight from an Anti to a Pro population.  
%C. The Schur decomposition of the weight matrix $W = V\Lambda V^{-1}$ is a unique decomposition with orthogonal $V$ and upper triangular $\Lambda$. Schur modes: $v_{\text{all}}$, $v_{\text{task}}$, $v_{\text{side}}$, and $v_{\text{diag}}$.  
%D. The marginal EPI distributions of the Schur eigenvalues at each level of task accuracy.
%E. The correlation of Schur eigenvalue with task performance in each learned EPI distribution.}
%\label{fig:SC}
%To quantify the emergent property of rapid task switching at various levels of accuracy, we considered the requirements of this model in this behavioral paradigm.
%At the end of successful trials, the response of the Pro population in the hemisphere of the correct choice must have a value near 1, while the Pro population in the opposite hemisphere must have a value near 0.
%Constraining a population response $x_\alpha \in \left[0,1\right]$ to be either 0 or 1 can be achieved by requiring that it has Bernoulli variance (see Section \ref{methods_SC}).
%Thus, we can formulate rapid task switching at a level of accuracy $p \in \left[0, 1\right]$ in both tasks in terms of the average steady response of the Pro population $\hat{p}$ of the correct choice,  the error in Bernoulli variance of that Pro neuron $\sigma^2_{\text{err}}$, and the average difference in Pro neuron responses $d$ in both Pro and Anti trials:
%\mathcal{B}(p) ~~\triangleq~~ \mathbb{E}\begin{bmatrix} \hat{p}_P \\ \hat{p}_A \\ (\hat{p}_P-p)^2 \\ (\hat{p}_A - p)^2 \\ \sigma^2_{P,err} \\ \sigma^2_{A,err} \\ d_P \\ d_A \end{bmatrix} = \begin{bmatrix} p \\ p \\ 0.15^2 \\ 0.15^2 \\ 0 \\ 0 \\ 1 \\ 1 \end{bmatrix}.
%Thus, $\mathcal{B}(p)$ denotes Bernoulli, winner-take-all responses between Pro neurons in a model executing rapid task switching near accuracy level $p$.
%We used EPI to learn distributions of the SC weight matrix parameters $z$ conditioned on of various levels of rapid task switching accuracy $\mathcal{B}(p)$ for $p \in \{50\%, 60\%, 70\%, 80\%, 90\%\}$.
%To make sense of these inferred distributions, we followed the approach of Duan et al. by decomposing the connectivity matrix $W = V\Lambda V^{-1}$ in such a way (the Schur decomposition) that the basis vectors $v_i$ are the same for all $W$ (Fig. \ref{fig:SC}C). 
%These basis vectors have intuitive roles in processing for this task, and are accordingly named the \textit{all} mode - all neurons co-fluctuate, \textit{side} mode - one side dominates the other, \textit{task} mode - the Pro or Anti populations dominate the other, and \textit{diag} mode - Pro- and Anti-populations of opposite hemispheres dominate the opposite pair. 
%The corresponding eigenvalues (e.g. $\lambda_{\text{task}}$, which change according to $W$) indicate the degree to which activity along that mode is increased or decreased by $W$. 
%We found that for greater task accuracies, the task mode eigenvalue increases, indicating the importance of $W$ to the task representation (Fig. \ref{fig:SC}D, purple; adjacent distributions from 60\% to 90\% have  $p<10^{-4}$, Mann-Whitney test with 50 estimates and 100 samples).
%Stepping from random chance (50\%) networks to marginally task-performing (60\%) networks, there is a marked decrease of the side mode eigenvalues (Fig. \ref{fig:SC}D, orange; $p<10^{-4}$).  
%Such side mode suppression relative to $50\%$ remains in the models achieving greater accuracy, revealing its importance towards task performance.
%There were no interesting trends with task accuracy in the all or diag mode (hence not shown in Fig. \ref{fig:SC}). 
%Importantly, we can conclude from our methodology that side mode suppression in $W$ allows rapid task switching, and that greater task-mode representations in $W$ increase accuracy.  
%These hypotheses are confirmed by forward simulation of the SC model (Fig. \ref{fig:SC}E, see Section \ref{methods_SC}) suggesting experimentally testable predictions: increase in rapid task switching performance should be correlated with changes in effective connectivity corresponding to an increase in task mode and decrease in side mode eigenvalues.

In a rapid task switching experiment \cite{duan2015requirement}, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from a visual stimulus in the Anti (A) task (Fig. \ref{fig:SC}A). 
Neural recordings in the midbrain superior colliculus (SC) exhibited two populations of neurons that simultaneously represented both task context (Pro or Anti) and motor response (contralateral or ipsilateral to the recorded side): the Pro/Contra and Anti/Ipsi neurons \cite{duan2018collicular}.
Duan et al. proposed a model of SC that, like the V1 model analyzed in the previous section, is a four-population dynamical system.  
We analyzed this model, where the neuron-type populations are functionally-defined as the Pro- and Anti-populations in each hemisphere (left (L) and right (R)), their connectivity is parameterized geometrically  (Fig. \ref{fig:SC}B).
The input-output function of this model is chosen such that the population responses $\mathbf{x} = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ are bounded from 0 to 1 as a function $f$ of a dynamically evolving internal variable $\mathbf{u}$.
The model responds to the side with greater Pro neuron activation; e.g. the reponse is left if $x_{LP} > x_{RP}$ at the end of the trial.
The dynamics evolve with timescale $\tau=0.09$ governed by connectivity weights $W$
\begin{equation}
\begin{split}
\tau \frac{d\mathbf{u}}{dt} &= -\mathbf{u} + W\mathbf{x} + \mathbf{h} + d\mathbf{B} \\
\mathbf{x} &= f(\mathbf{u})
\end{split}
\end{equation}
with white noise of variance $\epsilon^2 = 0.2^2$.
The input $\mathbf{h}$ is comprised of a cue-dependent input to the Pro or Anti populations, a stimulus orientation input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_SC}).
Here, we use EPI to determine the changes in network connectivity $\mathbf{z} = [sW, vW, dW, hW]^{\top}$ resulting in execution of rapid task switching behavior.

\begin{figure}
\begin{center}
\includegraphics[scale=0.48]{figures/fig4/fig4.pdf}
\end{center}
\caption{\small 
A. Rapid task switching behavioral paradigm (see text). 
B. Model of superior colliculus (SC). Neurons: LP - left pro, RP - right pro, LA - left anti, RA - right anti. 
Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
Subscripts $P$ and $A$ of connectivity weights indicate Pro or Anti populations.
C. The EPI parameter distribution of rapid task switching networks.  Black star indicates parameter choice u=of simulations.
D. Simulations of an SC network from the EPI distribution with 75\% accuracy in each task.  Top row shows no inactivation during Pro and Anti trials, and bottom row shows simulations with delay period inactivation (opto strength 0.7).  Shading indicates standard deviation across trials.
E. Difference in performance of each task during inactivation.  Inactivation level ``opto strength" scales from no inactivation (0) to full inactivation (1).  We compare delay period inactivation $1.2 < t < 1.5$ (blue), choice period inactivation $1.5 < t$ (red), and total inactivation $0 \leq t \leq 1.8$ (purple).
F. The effect of delay period inactivation on Anti accuracy versus dynamics eigenvalues.
}
\label{fig:SC}
\end{figure}

We define rapid task switching behavior as accurate execution of each task.  Inferred models should not exhibit fully random responses (50\%), or perfect performance (100\%), since perfection is never attained by even the best trained rats.
We formulate rapid task switching as an emergent property by stipulating that the average accuracy in the Pro task $p_P(\mathbf{x}, \mathbf{z})$ and Anti task $p_A(\mathbf{x}, \mathbf{z})$ be $75\%$ with variance $5\%^2$.
\begin{equation}\label{eq:EP}
\begin{split}
\mathcal{X} ~~:~~ \mathbb{E}_{\mathbf{z}}\begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  \begin{bmatrix} 75\% \\ 75\% \end{bmatrix}  \\ 
 \text{Var}_{\mathbf{z}}\begin{bmatrix} p_P(\mathbf{x}; \mathbf{z}) \\ p_A(\mathbf{x}; \mathbf{z}) \end{bmatrix}  &~~=~~  \begin{bmatrix} 5\%^2 \\ 5\%^2  \end{bmatrix}
\end{split}
\end{equation}
A variance of $5\%$ performance in each task will confer a posterior producing performances ranging from about $65\%-85\%$, allowing us to examine the properties of connectivity that yield better performance.

We ran EPI to obtain SC model connectivity parameters $z$ producing rapid task switching (Fig. \ref{fig:SC}C).
Some parameters were predictive of accuracy while others were not (Fig. \ref{fig:SC1}), and often had different effects on $p_P$ and $p_A$.
To make sense of this inferred distribution, we took the eigendecomposition of the symmetric connectivity matrices $W = V\Lambda V^{-1}$, which results in the same basis vectors $v_i$ for all $W$ parameterized by $z$ (Fig. \ref{fig:SC2}A). 
These basis vectors have intuitive roles in processing for this task, and are accordingly named the \textit{all} mode - all neurons co-fluctuate, \textit{side} mode - one side dominates the other, \textit{task} mode - the Pro or Anti populations dominate the other, and \textit{diag} mode - Pro- and Anti-populations of opposite hemispheres dominate the opposite pair.

Greater $\lambda_{\text{task}}$, $\lambda_{\text{side}}$, and $\lambda_{\text{diag}}$ all produce greater Pro accuracy.
This shows that strong task representations and hemispheral dominance in the dynamics result in better execution of the Pro task.
By visualizing these four variables together by $p_A$ (Fig. \ref{fig:SC3}B), we see that low $\lambda_{\text{task}}$ and $\lambda_{\text{diag}}$ producing strong Anti accuracy also have high $\lambda_{\text{side}}$ and $\lambda_{\text{all}}$.
Thus, stronger hemispheral dominance, relaxed task and diag mode dynamics, and slower circuit-wide decay result in greater Anti accuracy.

In agreement with experimental results form Duan et al., we found that inactivation above nominal strength during the delay period consistently decreased performance in the Anti task, but had no consistent effect on the Pro task (Fig. \ref{fig:SC}E) e.g. (Fig. \ref{fig:SC}D, bottom).
This difference in resiliency across tasks to delay perturbation is a prediction made by the inferred EPI distribution, rather than an emergent property that was conditioned upon.
Even though $p_P$ and $p_A$ are anticorrelated in the EPI posterior ($r=-0.15$, $p=3.68 \times 10^-12$), greater $p_P$ and $p_A$ both result in decreased resiliency to delay perturbation in the Anti task (Fig. \ref{fig:SC4}).
Ultimately, lower $\lambda_{\text{side}}$ and $\lambda_{\text{all}}$ and greater $\lambda_{\text{task}}$ produce networks more robust to delay perturbation (Fig. \ref{fig:SC}F)).

@@

\subsubsection{Example: 2D LDS}\label{methods_2DLDS}
To gain intuition for EPI, consider a two-dimensional linear dynamical system (2D LDS) model (Fig. S1A):
\begin{equation} 
\tau \frac{d\mathbf{x}}{dt} = A\mathbf{x}
\end{equation}
with
\begin{equation}
A = \begin{bmatrix} a_1 & a_2 \\ a_3 & a_4 \end{bmatrix}.
\end{equation}
To run EPI with the dynamics matrix elements as the free parameters $\mathbf{z} = [ a_1, a_2, a_3, a_4]$ (fixing $\tau=1$), the emergent property statistics $T(\mathbf{x})$ were chosen to contain the first and second moments of the oscillatory frequency, $\frac{\text{imag}(\lambda_1)}{2 \pi}$, and the growth/decay factor, $\text{real}(\lambda_1)$, of the oscillating system. 
 $\lambda_1$ is the eigenvalue of greatest real part when the imaginary component is zero, and alternatively of positive imaginary component when the eigenvalues are complex conjugate pairs.  
To learn the distribution of real entries of $A$ that produce a band of oscillating systems around 1Hz, we formalized this emergent property as $\text{real}(\lambda_1)$ having mean zero with variance $0.25^2$, and the oscillation frequency $2 \pi \text{imag}(\lambda_1)$ having mean $\omega = 1$ Hz with variance (0.1Hz)$^2$:
\begin{equation}
 \mathbb{E}\left[T(\mathbf{x}) \right] ~~ \triangleq ~~ \mathbb{E} \begin{bmatrix} \text{real}(\lambda_1) \\ \text{imag}(\lambda_1) \\ (\text{real}(\lambda_1)-0)^2  \\ (\text{imag}(\lambda_1)-2 \pi \omega)^2 \end{bmatrix} = \begin{bmatrix} 0.0 \\ 2 \pi \omega \\ 0.25^2 \\ (2 \pi 0.1)^2 \end{bmatrix} ~~ \triangleq ~~ \bm{\mu}.
 \end{equation} 
\begin{figure}
\begin{center}
\includegraphics[scale=0.58]{figures/figLDS1/figLDS1.pdf}
\end{center}
\begin{flushleft}
\caption{\small (LDS1): A. Two-dimensional linear dynamical system model, where real entries of the dynamics matrix $A$ are the parameters.  
B. The EPI distribution for a two-dimensional linear dynamical system with $\tau=1$ that produces an average of 1Hz oscillations with some small amount of variance.  Dashed lines indicate the parameter axes. 
C. Entropy throughout the optimization.  
At the beginning of each augmented Lagrangian epoch (2,000 iterations), the entropy dipped due to the shifted optimization manifold where emergent property constraint satisfaction is increasingly weighted.  
D. Emergent property moments throughout optimization.  
At the beginning of each augmented Lagrangian epoch, the emergent property moments adjust closer to their constraints.}
\end{flushleft}
\label{fig:LDS1}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{figures/figLDS2/figLDS2.pdf}
\end{center}
\begin{flushleft}
\caption{\small (LDS2): A. Probability contours in the $a_1$-$a_4$ plane were derived from the relationship to emergent property statistic of growth/decay factor $\text{real}(\lambda_1)$. 
B. Probability contours in the $a_2$-$a_3$ plane were derived from the emergent property statistic of oscillation frequency $2\pi \text{imag}(\lambda_1)$.}
\end{flushleft}
\label{fig:LDS2}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{figures/figLDS3/figLDS3.pdf}
\end{center}
\begin{flushleft}
\caption{\small (LDS3): Sampled dynamical systems $\mathbf{z} \sim q_{\bm{\theta}}(\mathbf{z})$ and their simulated activity from $\mathbf{x}(0) = [\frac{\sqrt{2}}{2}, -\frac{\sqrt{2}}{2}]$ colored by log probability. A. Each dimension of the simulated trajectories throughout time.  
B  The simulated trajectories in phase space.}
\end{flushleft}
\label{fig:LDS3}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.6]{figures/figSTG1/figSTG1.pdf}
\end{center}
\begin{flushleft}
\caption{\small (STG1): EPI optimization of the STG model producing network syncing. 
A. Entropy throughout optimization. 
B. The emergent property statistic means and variances converge to their constraints at 25,000 iterations following the fifth augmented Lagrangian epoch.}
\end{flushleft}
\label{fig:STG1}
\end{figure}

@@

%\begin{figure}
%\begin{center}
%\includegraphics[scale=0.5]{figures/figS5/figS5.pdf}
%\end{center}
%{\small Fig. S5: A. EPI for rank-1 networks doing noisy discrimination. B. EPI for rank-2 networks doing context-dependent discrimination.  See \cite{mastrogiuseppe2018linking} for theoretical equations and task description.}
%\end{figure}
